NOPT048 Notes

### 2019-07-18 misc/study-notes/ndmi011notes.tex
Added letters to theorems and lemmas

\proclaim Theorem. In any graph $G=(V,E)$,
$$\displaystyle\sum_{v\in V} \deg v = 2\cdot|E|$$

@@

\proclaim Theorem.
The number of vertices with odd degree is even.

@@

\proclaim Theorem.
A tree with $n$ vertices has $n-1$ edges.

@@

\thmparens{}{Generalised binomial theorem} For any $r\in \R$,
$$\sum_{i\geq 0} {r\choose i} x^i = (1 + x)^r.$$

@@

\proclaim Theorem. Let $(a_i)_{i=0}^\infty$ be a sequence of reals such that there exists some $k>0$ for which $|a_i|\leq k^i$ for every $i\geq 1$. Then for all $x$ in the interval $(-1/k, 1/k)$, the power series
$$a(x) = \sum_{i=0}^\infty a_ix^i$$
converges, so the generating function $a(x)$ has a closed form. Moreover, for any arbitrarily small $\epsilon >0$, the values of $a(x)$ for $x$ in the interval $(-\epsilon, \epsilon)$ uniquely determine the sequence $(a_i)_{i=0}^\infty$ where, by Taylor's theorem, $a_i = a^{(i)}(0)/i!$\slug

@@

\proclaim Lemma. For any s-t-flow $f$ and s-t-cut $\delta^+(S)$,
$$|f| = \sum_{e\in \delta^+(S)} f(e) - \sum_{e\in \delta^-(S)} f(e),$$
where $\delta^-(S)$ is the set $\{uv : u\notin S, v\in S\}$.

@@

\proclaim Theorem. If $f$ is a max-flow and $\delta^+(S)$ is a min-cut, then $|f| = c\big(\delta^+(S)\big)$.

@@

\proclaim Theorem. If for all $e\in E$, $c(e)\in \N_0$, then the capacity $n$ of the min-cut is integral and Algorithm M terminates in at most $n$ iterations.

@@

\proclaim Theorem. Let $f$ be a max-flow in a network with vertex capacities. If $T$ is a min-cut, then $|f| = d(T)$, and furthermore, if all capacities are integers, then there is an integer max-flow.

@@

\thmparens{}{Hall} A bipartite graph $G=(A\cup B, E)$ (with $A$ and $B$ disjoint) has a matching $M\subseteq E$ of size $|M| = |A|$ if and only if $|N(S)|\geq |S|$ for all $S\subseteq A$.

@@

\thmparens{}{K\H onig's theorem} Let $G$ be a bipartite graph with maximum matching $M$ and minimum vertex cover $C$. Then $|M| = |C|$.\slug

@@

\proclaim Lemma. Let $G=(V,E)$ be a graph. Then for all edges $uv\in E$, $\kappa(G\setminus\{uv\})\geq \kappa(G) - 1$.

@@

\proclaim Lemma. Let $G=(V,E)$ be a graph with $|V|=n$. Then the following both hold:
\medskip
\item {1.} $\kappa(G) - 1 \leq \kappa(G-v)$ for all $v\in V$.
\smallskip
\item {2.}$\lambda(G) - 1 \leq \lambda(G-e)\leq \lambda(G)$ for all $e\in E$.
\medskip

@@

\proclaim Lemma. Let $G=(V,E)$ be a graph and let $\delta(G)$ denote the minimum degree over all vertices in $G$. Then
$$ \kappa(G) \leq \lambda(G) \leq \delta(G).$$

@@

\proclaim Theorem. A graph is 2-connected if and only if it has an ear decomposition.

@@

\proclaim Theorem. Let $e$ be an edge in $K_n$. Then
$$ T(K_n-e) = (n-2)n^{n-3}.$$

@@

\proclaim Theorem. The maximum number of edges of any graph $G$ not containing $C_3$ as a subgraph is at most $\lfloor n^2/4 \rfloor$.

@@

\proclaim Theorem. The maximum number of edges of any graph $G$ not containing $C_4$ as a subgraph is at most $(n^{3/2} + n) / 2$.

@@

\proclaim Lemma. Let $(X,{\cal L})$ be a finite projective plane. Then for all $L, L'\in {\cal L}$, there exists an $z\in L$ such that $z\notin L\cup L'$

@@

\proclaim Lemma. Let $(X,{\cal L})$ be a finite projective plane. Then for all $L,L'\in {\cal L}$, $|L| = |L'|$.

@@

\proclaim Theorem. Let $(X,{\cal L})$ be a finite projective plane of order $n$. Then the following statements all hold:
\medskip
\item {1.} Exactly $n+1$ lines pass through any point $x\in X$.
\smallskip
\item {2.} $|X| = n^2 + n + 1$.
\smallskip
\item {3.} $|{\cal L}| = n^2 + n + 1$.
\medskip

@@

\proclaim Lemma. The dual $({\cal L}, \Lambda)$ of a finite projective plane $(X,{\cal L})$ is also a finite projective plane.

@@

\proclaim Theorem. A finite projective plane exists of order $n$ if $n$ is a prime power.

@@

\proclaim Theorem. Let $M$ be a set of pairwise orthogonal Latin squares of order $n$. Then $|M|\leq n-1$.

@@

\proclaim Theorem.
    For $n\geq 2$, a finite projective plane of order $n$ exists if and only if there exists a set of $n-1$ pairwise orthogonal Latin squares of order $n$.

@@

\proclaim Theorem. Let $G=(V,E)$ be a graph. If $|V|\geq {k-l-2\choose k-1} = {k+l-2\choose l-1}$, then $\alpha(G)\geq l$ or $\omega(G)\geq k$.

@@

\proclaim Theorem. For $k\geq 3$, $r(k,k)>2^{k/2}$.

@@

\proclaim Theorem. For a linear code $C$,
$$ \Delta(C) = \min\{\hbox{wt}\,x : x\in C\ \hbox{and}\ x\neq 0\}.$$

@@

\thmparens{S}{Hamming bound} If an $(n,k,d)_q$-code exists, then
$$q^k\cdot\hbox{\rm vol}\left(\left\lfloor {d-1\over 2} \right\rfloor, n\right) \leq q^n.$$

@@

\thmparens{B}{Singleton bound} If $C$ is an $(n,k,d)_q$-code, then $k\leq n-d+1$.

>>>>

\proclaim Theorem D. In any graph $G=(V,E)$,
$$\displaystyle\sum_{v\in V} \deg v = 2\cdot|E|$$

@@

\proclaim Theorem E. The number of vertices with odd degree is even.

@@

\proclaim Theorem T.
A tree with $n$ vertices has $n-1$ edges.

@@

\thmparens{G}{Generalised binomial theorem} For any $r\in \R$,
$$\sum_{i\geq 0} {r\choose i} x^i = (1 + x)^r.$$

@@

\proclaim Theorem S. Let $(a_i)_{i=0}^\infty$ be a sequence of reals such that there exists some $k>0$ for which $|a_i|\leq k^i$ for every $i\geq 1$. Then for all $x$ in the interval $(-1/k, 1/k)$, the power series
$$a(x) = \sum_{i=0}^\infty a_ix^i$$
converges, so the generating function $a(x)$ has a closed form. Moreover, for any arbitrarily small $\epsilon >0$, the values of $a(x)$ for $x$ in the interval $(-\epsilon, \epsilon)$ uniquely determine the sequence $(a_i)_{i=0}^\infty$ where, by Taylor's theorem, $a_i = a^{(i)}(0)/i!$\slug

@@

\proclaim Lemma D. For any s-t-flow $f$ and s-t-cut $\delta^+(S)$,
$$|f| = \sum_{e\in \delta^+(S)} f(e) - \sum_{e\in \delta^-(S)} f(e),$$
where $\delta^-(S)$ is the set $\{uv : u\notin S, v\in S\}$.

@@

\proclaim Theorem C. If $f$ is a max-flow and $\delta^+(S)$ is a min-cut, then $|f| = c\big(\delta^+(S)\big)$.

@@

\proclaim Theorem I. If for all $e\in E$, $c(e)\in \N_0$, then the capacity $n$ of the min-cut is integral and Algorithm M terminates in at most $n$ iterations.

@@

\proclaim Theorem J. Let $f$ be a max-flow in a network with vertex capacities. If $T$ is a min-cut, then $|f| = d(T)$, and furthermore, if all capacities are integers, then there is an integer max-flow.

@@

\thmparens{H}{Hall} A bipartite graph $G=(A\cup B, E)$ (with $A$ and $B$ disjoint) has a matching $M\subseteq E$ of size $|M| = |A|$ if and only if $|N(S)|\geq |S|$ for all $S\subseteq A$.

@@

\thmparens{K}{K\H onig's theorem} Let $G$ be a bipartite graph with maximum matching $M$ and minimum vertex cover $C$. Then $|M| = |C|$.\slug

@@

\proclaim Lemma K. Let $G=(V,E)$ be a graph. Then for all edges $uv\in E$, $\kappa(G\setminus\{uv\})\geq \kappa(G) - 1$.

@@

\proclaim Lemma R. Let $G=(V,E)$ be a graph with $|V|=n$. Then the following both hold:
\medskip
\item {1.} $\kappa(G) - 1 \leq \kappa(G-v)$ for all $v\in V$.
\smallskip
\item {2.}$\lambda(G) - 1 \leq \lambda(G-e)\leq \lambda(G)$ for all $e\in E$.
\medskip

@@

\proclaim Lemma C. Let $G=(V,E)$ be a graph and let $\delta(G)$ denote the minimum degree over all vertices in $G$. Then
$$ \kappa(G) \leq \lambda(G) \leq \delta(G).$$

@@

\proclaim Theorem E. A graph is 2-connected if and only if it has an ear decomposition.

@@

\proclaim Theorem L. Let $e$ be an edge in $K_n$. Then
$$ T(K_n-e) = (n-2)n^{n-3}.$$

@@

\proclaim Theorem T. The maximum number of edges of any graph $G$ not containing $C_3$ as a subgraph is at most $\lfloor n^2/4 \rfloor$.

@@

\proclaim Theorem F. The maximum number of edges of any graph $G$ not containing $C_4$ as a subgraph is at most $(n^{3/2} + n) / 2$.

@@

\proclaim Lemma Z. Let $(X,{\cal L})$ be a finite projective plane. Then for all $L, L'\in {\cal L}$, there exists an $z\in L$ such that $z\notin L\cup L'$

@@

\proclaim Lemma C. Let $(X,{\cal L})$ be a finite projective plane. Then for all $L,L'\in {\cal L}$, $|L| = |L'|$.

@@

\proclaim Theorem F. Let $(X,{\cal L})$ be a finite projective plane of order $n$. Then the following statements all hold:
\medskip
\item {1.} Exactly $n+1$ lines pass through any point $x\in X$.
\smallskip
\item {2.} $|X| = n^2 + n + 1$.
\smallskip
\item {3.} $|{\cal L}| = n^2 + n + 1$.
\medskip

@@

\proclaim Lemma D. The dual $({\cal L}, \Lambda)$ of a finite projective plane $(X,{\cal L})$ is also a finite projective plane.

@@

\proclaim Theorem P. A finite projective plane exists of order $n$ if $n$ is a prime power.

@@

\proclaim Theorem L. Let $M$ be a set of pairwise orthogonal Latin squares of order $n$. Then $|M|\leq n-1$.

@@

\proclaim Theorem O.
    For $n\geq 2$, a finite projective plane of order $n$ exists if and only if there exists a set of $n-1$ pairwise orthogonal Latin squares of order $n$.

@@

\proclaim Theorem R. Let $G=(V,E)$ be a graph. If $|V|\geq {k-l-2\choose k-1} = {k+l-2\choose l-1}$, then $\alpha(G)\geq l$ or $\omega(G)\geq k$.

@@

\proclaim Theorem L. For $k\geq 3$, $r(k,k)>2^{k/2}$.

@@

\proclaim Theorem L. For a linear code $C$,
$$ \Delta(C) = \min\{\hbox{wt}\,x : x\in C\ \hbox{and}\ x\neq 0\}.$$

@@

\thmparens{H}{Hamming bound} If an $(n,k,d)_q$-code exists, then
$$q^k\cdot\hbox{\rm vol}\left(\left\lfloor {d-1\over 2} \right\rfloor, n\right) \leq q^n.$$

@@

\thmparens{S}{Singleton bound} If $C$ is an $(n,k,d)_q$-code, then $k\leq n-d+1$.

### 2019-07-18 misc/study-notes/nopt048notes.tex
Fixed errors in both notes
Added letters to theorems and lemmas

\maketitle{NOPT048 Linear Programming and Combinatorial Optimization}{Notes by}{Marcel K. Goh {\rm (Prague, Czech Rep.)}}{9 June 2019}

@@

A polyhedron is {\it pointed} if it does not contain a line.
\proclaim Lemma. Every non-empty pointed polyhedron $P$ has at least one vertex.

\proclaim Theorem. Consider the linear program
$$\matrix{
    & \hbox{\rm maximise}\hfill & \quad & c^Tx \hcr
    & \hbox{\rm subject to}\hfill & & Ax \leq b.\hfill
}$$
If the polyhedron $P$ defined by $Ax\leq b$ is pointed and if an optimum exists, then there exists a vertex of $P$ that attains the maximum.

@@

\proclaim Theorem. Let $M=(E,I)$ be a matroid with rank function $r$. The weighted greedy algorithm finds the maximum weighted independent set $S_k$ of cardinality $k$ for each $1\leq k\leq r(E)$.\slug

\thmparens{G}{Greedy equivalence} Let $M=(E,I)$ be an independence system. Then if the weighted greedy algorithm works for all (non-negative) weights, then $M$ is a matroid.\slug

>>>>

\maketitle{NOPT048 Linear Programming and Combinatorial Optimization\footnote{$^*$}{\eightpt Course given by Prof. Hans Raj Tiwary at Charles University in Prague}}{Notes by}{Marcel K. Goh {\rm (Prague, Czech Rep.)}}{9 June 2019}

@@

A polyhedron is {\it pointed} if it does not contain a line.
\proclaim Lemma P. Every non-empty pointed polyhedron $P$ has at least one vertex.

\proclaim Theorem E. Consider the linear program
$$\matrix{
    & \hbox{\rm maximise}\hfill & \quad & c^Tx \hcr
    & \hbox{\rm subject to}\hfill & & Ax \leq b.\hfill
}$$
If the polyhedron $P$ defined by $Ax\leq b$ is pointed and if an optimum exists, then there exists a vertex of $P$ that attains the maximum.

@@

\proclaim Theorem G. Let $M=(E,I)$ be a matroid with rank function $r$. The weighted greedy algorithm finds the maximum weighted independent set $S_k$ of cardinality $k$ for each $1\leq k\leq r(E)$.\slug

\thmparens{E}{Greedy equivalence} Let $M=(E,I)$ be an independence system. Then if the weighted greedy algorithm works for all (non-negative) weights, then $M$ is a matroid.\slug
